Loss recovery and tail loss performance in Multipath TCP

### 2016-10-05 main.tex
initial merge of retrans and tlp

\documentclass[12pt,draftcls,onecolumn]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[dvips]{graphicx}
% Title Page
\title{Understanding Loss recovery in MultiPath TCP }
\author{KAU}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.9\textwidth]{images/fortest.ps}
\caption{Topology used for Emulation}\label{fig1}
\end{center}
\label{prl}
\end{figure}
\begin{center}
\begin{table}
\begin{center}
\begin{tabular}{|c|cccccccccc|}
      \hline
      \multicolumn{1}{c}{} & & \\[\dimexpr-\normalbaselineskip-\arrayrulewidth]
      \textbf{Burst Size} & \multicolumn{10}{c|}{7 Packets} \\
      \hline
      \textbf{Separation Time} & \multicolumn{10}{c|}{2s} \\
      \hline

@@

\section{Observations and Discussion}\label{disc}
Preliminary experiments on CORE emulator with the setup mentioned in section\ref{exsetup} showed several cases where the behavior on losses differ.
These cases can be categorized as 
\begin{itemize}
 \item SYN Loss
 \item First packet loss (in second subflow)
 \item RTO 
 \item Unwanted RST 
 \item Tail loss
 \item Specific packet loss - in meta level and mptcp level

\end{itemize}

xplot analysis indicates that SYN loss triggers an RTO in the worst case and incurring a latency of 3 RTTs at the sender.

@@

\begin{table}[]
\centering
\caption{Tail loss scenarios with tcp.early.retrans = 3 default}
\label{ret3}
\begin{tabular}{|l|l|l|l|l|}
\hline
 testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \hline
\end{tabular}
\end{table}

@@

>>>>

\documentclass[12pt,draftcls,onecolumn]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[dvips]{graphicx}
% Title Page
\title{Loss recovery and tail loss performance in Multipath TCP}
\author{KAU}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.9\textwidth]{images/fortest.ps}
\caption{Topology used for Emulation}\label{fig1}
\end{center}
\label{prl}
\end{figure}
\begin{center}
\begin{table}
\begin{center}
\begin{tabular}{|c|cccccccccc|}
      \hline
      \multicolumn{1}{c}{} & & \\[\dimexpr-\normalbaselineskip-\arrayrulewidth]
      \textbf{Burst Size} & \multicolumn{10}{c|}{160 Packets} \\
      \hline
      \textbf{Separation Time} & \multicolumn{10}{c|}{2s} \\
      \hline

@@

\subsection{Testing with Deterministic loss patterns}
In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.

\section{Observations and Discussion}\label{disc}
Preliminary experiments on CORE emulator with the setup mentioned in section\ref{exsetup} showed several cases where the behavior on losses differ.

@@

\begin{table}[]
\centering
\caption{Tail loss scenarios with tcp.early.retrans = 3 default}
\label{ret3}
\begin{tabular}{|l|l|l|l|l|}
\hline
 testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                &  \\ \hline
\end{tabular}
\end{table}

\begin{table}[]
\centering
\caption{Tail loss scenarios with tcp.early.retrans = 0 ER disabled}
\label{ret3}
\begin{tabular}{|l|l|l|l|l|}
\hline
 testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
Tail drop 4 & 20ms - 120ms &  			&		&  \\ \hline
Tail drop 5 & 120ms - 20ms &  			& 		& \\ \hline 
\end{tabular}
\end{table}

\begin{table}[]
\centering
\caption{Tail loss scenarios with tcp.early.retrans = 4 TLP disabled}
\label{ret3}
\begin{tabular}{|l|l|l|l|l|}
\hline
 testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
\end{tabular}
\end{table}

@@

\section{Improvements to TLP}
The goal of Tail Loss Probe(TLP) is to reduce tail latency of short flows. It achieves this by converting retransmission timeouts (RTOs) occuring due to tail losses (losses at end of transactions) into fast recovery. TLP transmits one packet in two round-trips when a connection is in Open state and isn't receiving any ACKs. The transmitted packet, aka loss probe, can be either new or a retransmission. When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout.

Current implementation of TLP is at the TCP flow level. In the event of RTO timeout the MPTCP retransmission happens with a reinjection in to scheduler along with sending lost packet on the same path. But in the Event of Probe timeout, the lost packet is being sent on the same path without injecting in to scheduler.  Modifications should reinject the lost packet in to the mptcp scheduler in the event of PTO.

\subsection{Example scenario where MPTCP TLP could incur more latency}

Use a traffic pattern where there is new burst to send before the RTO after a tail loss triggering TLP on a flow.

\section{Proposed changes and expected implications}

\subsection{Implementation Plan}

Things to do:

When PTO fires: in code trigger call on PTO in tcp level
It calls ${tcp\_push\_one()}$ in tcp-output.c and then ${mptcp\_write\_xmit()}$.
As it is reinjection (due to only one case currently in MPTCP), that packet will have path mask
to put it on the same subflow. Here we need to add a case for reinjection due to PTO that should
send the packet to $reinject\_queue$ of ${mptcp\_cb}$ as well.

Main flow packet queue and subflow packet queue: ${meta\_sk}$ refers to the
main mptcp flow.

Loss probe timer used in Linux: ${ICSK\_TIME\_LOSS\_PROBE}$.

\section{Evaluation of proposal}

Initial testing with customized program to send bursts of packets with fixed burst seperation. Adjust the burst seperation to get the desired scenario.
